Take A Relaxing Bath

### 2016-02-02 Tex/paper.tex
final correction added

\title{Human Capital Model} \date{\today{}}
\maketitle

@@

\subsection{Restatement of the Problem}
\label{sec:restatement-of-the-problem}
We are required to build a mathematical model to solve the problem of heating water in bathtub. Thus we have some subproblems:
\begin{itemize}
\item Build a basic model that demonstrates the change of the temperature in the bathtub in space and time without other interventions.
\item Figure out the influence of parameters such as the shape of the bathtub or the motions made by the person or so.
\item Propose a strategy to keep the temperature as even as possible.
\end{itemize}
In the first step, we build the simplest model. The shape of the bathtub is a cuboid, and the person will stay still. In the second step, the person moves slowly and the shape of the bathtub changes. In the third step, we develop the conclusion of the strategy.

@@

Definitions of symbols employed in this paper are listed in Table \ref{1_t} and Table \ref{2_t}.
\begin{table}
\begin{tabular}{l|l}
  Constant & Description \\
  \hline
  $k$            &Thermal conductivity of water \\
  $\rho$         &Density of water \\
  ${\rho}_a$     &Density of air \\
  ${\gamma}_1$   &Heat exchange coefficient between water and air \\
  ${\gamma}_2$   &Heat exchange coefficient between water and person \\
  ${\gamma}_3$   &Heat exchange coefficient between water and bathtub \\
  $C_w$          &Specific heat capacity of water \\
  $C_a$          &Specific heat capacity of air \\
  $V$            &Total volume of water \\
  $V_a$          &Total volume of air above the water \\
  $v$            &Flow velocity of air\\
  $x_s$          &Humidity ratio in saturated air \\
  $x$            &Humidity ratio in the air \\
  $x_0$          &Initial humidity ratio in the air \\
  $A$            &Square of water surface \\
  $V_a$          &Volume of surrounding air \\
  $C_v$          &Evaporation heat of water \\
\end{tabular}
\caption{Constant}\label{1_t}
\end{table}

\begin{table}
\begin{tabular}{l|l}
  Variable & Description \\
  \hline
  $t$            &time \\
  $u$            &Temperature of water \\
  $u$            &Temperature of water when t=0 \\
  $u_1$          &Temperature of air \\
  $u_2$          &Temperature of person's surface \\
  $u_3$          &Temperature of bathtub \\
  $u_h$          &Temperature of hot water from faucet \\
  $v_h$          &Velocity of hot water from faucet \\
  $g$            &Evaporation heat of water\\
\end{tabular}
\caption{Variable}\label{2_t}
\end{table}

@@

\begin{itemize}
\item \textbf{assumption 1} The surface temperature of bathtub and the person keeps constant.

@@

\item \textbf{assumption 5} Water, bathtub, person and the surrounding air compose a closed heat system.
\end{itemize}

% the main section of the paper
\section{Water in the Bathtub Model}
\label{sec:human-capital-model}

@@

In our model, five factors account for heat transfer of water: heat exchange between water and air, heat
exchange between water and the person, heat exchange between water and the bathtub, heat energy of hot water from the
faucet and heat energy of water flow outside. According to thermology, we deduce the equation
of water temperature in the bathtub. The initial-boundary value conditions are deduced according to
Fourier theorem, vaporization equation\cite{1}.

@@

Figure \ref{1_p} shows thermal analysis of water in bathtub
In the photo, $W_1$ is heat exchange between the person and water, while $W_2$ is heat from water vaporization. $W_3$ stands for heat exchange between the bathtub and water, and $W_4$ is heat conductivity between water and air. $W_5$ represents heat exchange between the bathtub and air, and $W_6$ is heat increasement with hot in-water.

@@

First, we assume that the heat change of all the water in a tiny time $\Delta t$ from $t_1$ to $t_2fv$ is $\Delta Q$.
As shown in Figure 1, heat of water in the bathtub exchanges with air, the person and the bathtub.
In addition, heat changes when the hot water flows in and the excess water flows out.
Thus $\Delta Q$ is divided into five parts, i.e.
\begin{equation}
 \Delta Q=\Delta Q_1+\Delta Q_2-\Delta Q_3-\Delta Q_4
\end{equation}
in which $\Delta Q_1$ is the heat transferred from air, person and bathtub,
$\Delta Q_2$ is the heat of the hot water from the faucet in $\Delta t$,
$\Delta Q_3$ is the heat of the water flows away in $\Delta t$,
$\Delta Q_4$ is the heat loss because of vaporization.

@@

The boundary of water consists of four parts: water-air boundary, water-person boundary, water-bathtub boundary, and the square for faucet. We denote them as
${\Gamma}_1$, ${\Gamma}_2$, ${\Gamma}_3$, ${\Gamma}_4$ respectively.

Since bather is temperature-constant, the boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered
as the Dirichlet value condition. That is
\begin{equation}
 u=u_i, (x,y,z)\in {\Gamma}_i, i=2.4
\end{equation}
In the equation, $u_2$ is the surface temperature of the person. We assume it is constant. $u_4$ is the temperature of hot water.

@@

$u_1$, the temperature of air, is time-dependant, because water vapor transfer heat to air. We can calculate $u_1$ by calculating the energy of water vapor evaporating in the air. Referencing the deduction of evaporation in the last section,
\begin{equation}
  u_1=u_0+\frac{C_w\rho}{C_a{{\rho}_a}} (x_s-x_0)(1-e^{-\lambda t})
\end{equation}

@@

\subsection{PDE Solution}
\label{sec:PDE solution}
Since our equations are complex and the region is quite irregular, it's impossible to get a analytical solution. Finite element method is normally adopted to deal with this kind of problem.

@@

In each operation, $u_2$ is taken as $33.5\,^{\circ}\mathrm{C}$
and the initial value condition is $35\,^{\circ}\mathrm{C}$.
 We take 120s as time period. The following graphs show the results we get.

@@

\subsection{Parameters of Person}
\label{sec:parameters of person}
\begin{itemize}
\item Comparing the temperature graphs with different shapes of the bather, we find when the shape gets more regular, $u_h$, $v_h$ decreases relatively.
\item Change the volume of the bather. The result shows that, $u_h$, $v_h$ for best strategy increases when the volume increases.
\item Change the surface temperature of the bather to get the best strategy. The result shows no noticeable change.
\begin{table}
\begin{center}
\begin{tabular}{c|c|c}
\hline
 Surface Temperature     &$v_h$  &$u_h$        \\ \hline
 $u_2=33.5$    &0.2             & 40              \\ \hline
 $u_2=33.5$    &0.15            & 40              \\ \hline
 $u_2=33.5$  &0.2             & 45              \\ \hline
\end{tabular}
\end{center}
\caption{Surface Temperature}\label{5_t}
\end{table}
\end{itemize}

@@

\section{Non-technical explanation}

>>>>

\title{Take A Relaxing Bath} \date{\today{}}
\maketitle

@@

\subsection{Restatement of the Problem}
\label{sec:restatement-of-the-problem}
We are required to build a mathematical model to solve the problem of heating water in bathtub. Thus we have some subproblems:
\begin{itemize}
\item Build a basic model that demonstrates the change of the temperature in the bathtub in space and time without other interventions.
\item Figure out the influence of parameters such as the shape of the bathtub or the motions made by the bather or so.
\item Propose a strategy to keep the temperature as even as possible.
\end{itemize}
In the first step, we build the simplest model. The shape of the bathtub is a cuboid, and the bather will stay still. In the second step, the bather moves slowly and the shape of the bathtub changes. In the third step, we develop the conclusion of the strategy.

@@

Definitions of symbols employed in this paper are listed in Table \ref{1_t} and Table \ref{2_t}.
\begin{table}
\begin{tabular}{l|l}
  Constant & Description \\
  \hline
  $k$            &Thermal conductivity of water \\
  $\rho$         &Density of water \\
  ${\rho}_a$     &Density of air \\
  ${\gamma}_1$   &Heat exchange coefficient between water and air \\
  ${\gamma}_2$   &Heat exchange coefficient between water and bather \\
  ${\gamma}_3$   &Heat exchange coefficient between water and bathtub \\
  $C_w$          &Specific heat capacity of water \\
  $C_a$          &Specific heat capacity of air \\
  $V$            &Total volume of water \\
  $V_a$          &Total volume of air above the water \\
  $v$            &Flow velocity of air\\
  $x_s$          &Humidity ratio in saturated air \\
  $x$            &Humidity ratio in the air \\
  $x_0$          &Initial humidity ratio in the air \\
  $A$            &Square of water surface \\
  $V_a$          &Volume of surrounding air \\
  $C_v$          &Evaporation heat of water \\
\end{tabular}
\caption{Constant}\label{1_t}
\end{table}

\begin{table}
\begin{tabular}{l|l}
  Variable & Description \\
  \hline
  $t$            &time \\
  $u$            &Temperature of water \\
  $u$            &Temperature of water when t=0 \\
  $u_1$          &Temperature of air \\
  $u_2$          &Temperature of bather's surface \\
  $u_3$          &Temperature of bathtub \\
  $u_h$          &Temperature of hot water from faucet \\
  $v_h$          &Velocity of hot water from faucet \\
  $g$            &Evaporation heat of water\\
\end{tabular}
\caption{Variable}\label{2_t}
\end{table}

@@

\begin{itemize}
\item \textbf{assumption 1} The surface temperature of bathtub and the bather keeps constant.

@@

\item \textbf{assumption 5} Water, bathtub, bather and the surrounding air compose a closed heat system.
\end{itemize}

% the main section of the paper
\section{PDE Model for Heating Water}
\label{sec:human-capital-model}

@@

In our model, five factors account for heat transfer of water: heat exchange between water and air, heat
exchange between water and the bather, heat exchange between water and the bathtub, heat energy of hot water from the
faucet and heat energy of water flow outside. According to thermology, we deduce the equation
of water temperature in the bathtub. The initial-boundary value conditions are deduced according to
Fourier theorem, vaporization equation\cite{1}.

@@

Figure \ref{1_p} shows thermal analysis of water in bathtub
In the photo, $W_1$ is heat exchange between the bather and water, while $W_2$ is heat from water vaporization. $W_3$ stands for heat exchange between the bathtub and water, and $W_4$ is heat conductivity between water and air. $W_5$ represents heat exchange between the bathtub and air, and $W_6$ is heat increasement with hot in-water.

@@

First, we assume that the heat change of all the water in a tiny time $\Delta t$ from $t_1$ to $t_2fv$ is $\Delta Q$.
As shown in Figure 1, heat of water in the bathtub exchanges with air, the bather and the bathtub.
In addition, heat changes when the hot water flows in and the excess water flows out.
Thus $\Delta Q$ is divided into five parts, i.e.
\begin{equation}
 \Delta Q=\Delta Q_1+\Delta Q_2-\Delta Q_3-\Delta Q_4
\end{equation}
in which $\Delta Q_1$ is the heat transferred from air, bather and bathtub,
$\Delta Q_2$ is the heat of the hot water from the faucet in $\Delta t$,
$\Delta Q_3$ is the heat of the water flows away in $\Delta t$,
$\Delta Q_4$ is the heat loss because of vaporization.

@@

The boundary of water consists of four parts: water-air boundary, water-bather boundary, water-bathtub boundary, and the square for faucet. We denote them as
${\Gamma}_1$, ${\Gamma}_2$, ${\Gamma}_3$, ${\Gamma}_4$ respectively.

Since bather is temperature-constant, the boundary value condition of ${\Gamma}_1$ and ${\Gamma}_3$ could be considered
as the Dirichlet value condition. That is
\begin{equation}
 u=u_i, (x,y,z)\in {\Gamma}_i, i=2.4
\end{equation}
In the equation, $u_2$ is the surface temperature of the bather. We assume it is constant. $u_4$ is the temperature of hot water.

@@

$u_1$, the temperature of air, is time-dependant because water vapor transfer heat to air. We can calculate $u_1$ by calculating the energy of water vapor evaporating in the air. Referencing the deduction of evaporation in the last section,
\begin{equation}
  u_1=u_0+\frac{C_w\rho}{C_a{{\rho}_a}} (x_s-x_0)(1-e^{-\lambda t})
\end{equation}

@@

\subsection{PDE Solution}
\label{sec:PDE solution}
Since our equations are complex and the region is quite irregular, it's impossible to get an analytical solution. Finite element method is normally adopted to deal with this kind of problem.

@@

In each operation, $u_2$ is taken as $33.5\,^{\circ}\mathrm{C}$
and the initial value condition is $35\,^{\circ}\mathrm{C}$.
 We set 120s as time period. The following graphs show the results we get.

@@

\subsection{Parameters of Bather}
\label{sec:parameters of bather}
\begin{itemize}
\item Comparing the temperature graphs with different shapes of the bather, we find when the shape gets more regular, $u_h$, $v_h$ decreases relatively.
\item Change the volume of the bather. The result shows that, $u_h$, $v_h$ for best strategy increases when the volume increases.
\item Change the surface temperature of the bather to get the best strategy. The result shows no noticeable change.
\begin{table}
\begin{center}
\begin{tabular}{c|c|c}
\hline
 Surface Temperature     &$v_h$  &$u_h$        \\ \hline
 $u_2=33.5$    &0.2             & 40              \\ \hline
 $u_2=33.5$    &0.15            & 40              \\ \hline
 $u_2=33.5$  &0.2             & 45              \\ \hline
\end{tabular}
\end{center}
\caption{Surface Temperature}\label{5_t}
\end{table}
\end{itemize}

@@

\section{Non-technical Explanation}
